---
title: "Lab#1"
author: "Tyler Cobian"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)

```
Read in the data file for the NOAA fisheries
```{r, echo=FALSE, include=FALSE}
us_landings<- read_csv(here("data", "noaa_fisheries.csv"))

```
Now make a tidy version of the data frame
```{r}
landings_tidy<- us_landings %>%
  janitor::clean_names() %>%# the :: just states which package the function is in 
  mutate(state = str_to_lower(state),# naming the new column the same name of an existing colum, will override the old column
         afs_name = str_to_lower(afs_name)) %>% 
  mutate(dollars_usd = parse_number(dollars_usd))# parse_number will drop and none numeric character from a number
```
Let's get some info for only salmon
```{r}
salmon_landings<- landings_tidy %>% 
  mutate(afs_clean = str_remove(afs_name, pattern = "aggregate")) %>% # got ride of any "aggregate"
  filter(str_detect(afs_clean, pattern = "salmon")) %>% # only keep anything with the name salmon
  separate(afs_clean, into = c("group", "species"), sep = ", ") #breaks up the group and species of the fish into different columns, "sep = " tells R where to seperate the str
 
```
Find some grouped summary data:
Find annual total US landings and dollar value (summing across all states) for each type of salmon using group_by() and summarise() functions
```{r}
salmon_summary<- salmon_landings %>% 
  group_by(year, species) %>% 
  summarize(
    tot_landings = sum(landings_pounds),
    tot_value = sum(dollars_usd)
  )

```
Now we will make a graph for this data
```{r}
salmon_landings_graph<- ggplot(data = salmon_summary, aes(x = year, y = tot_landings))+
  geom_line(aes(color = species))+
  theme_get()
salmon_landings_graph
```
Now we will export our graph with ggsave
```{r}
ggsave(plot = salmon_landings_graph, 
       here("figures", "Us_Sallmon_TC.jpg"), 
       height = 5,
       width = 8)
# ggsave will save export and image
# here() will twll it where to save the image
# you can also manipulate the type of file to save it as and  the height and width

```
Make a nice kable table
```{r}
salmon_first_5<- salmon_summary %>% 
  head(5)
salmon_first_5
```
Use kable to make a better table
```{r}
kable(salmon_first_5) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```























